Rule Synthesis for Rule-based Intrusion Detectors

### 2020-05-19 paper/paper.tex
opening space for Lucas/Guilherme

@@

Our experiments is affected by the usual threats. Considering external
validity. Our results cannot be generalized beyond the subjects we
selected. To mitigate that threat we selected a variety of attacks
with different characteristics. Considering internal validity, the
implementation of \tname\ could contain errors. We mitigated that
threat by carefully inspecting code and results. Considering construct
validity, the choice of metric used to evaluate the approach may not
be applicable in practice. We considering the use of ranking of the
golden rule a reasonable choice as the approach places the golden rule
close to the top in most cases.

\section{Discussion of other attacks}
\label{sec:rce}
\label{sec:jboss}
\label{sec:content-example}
\label{sec:dos}

%\Mar{@Lucas, por
%  favor, observe o padrao usado nos exemplos anteriores. eh preciso
%  explicar em alto nivel o ataque antes de explicar a regra!}

%% In an RCE attack, hackers
%% intentionally exploit a vulnerability in the system to run malware
%% By prompting the targeted device to perform code execution, a hacker
%% can run arbritrary code to
%% their own programming in its
%% place. This programming can then enable them to gain full access,
%% steal data, carry out a full distributed denial of service (DDoS)
%% attack, destroy files and infrastructure, or engage in illegal
%% activity. And as the term remote execution suggests, an RCE
%% cyberattack can take place from any geophysical location."\\

%A opção  define que o match só será feito em pacotes que fazem parte
%de uma conexão já estabelecida entre um cliente e um servidor. Além
%disso, define também que os pacotes devem estar direcionados ao servidor, não ao cliente.

\begin{figure}[t!]
%  \vspace{-1ex}    
  \lstinputlisting[language=C,numbers=none,keywords={content,flow}]{adaptor-golden-rule.suricata}
%  \vspace{-2ex}  
  \caption{\label{fig:rule-jboss}Suricata rule for JBoss Remote Code
    Execution.}
%  \vspace{-1ex}
\end{figure}

\Mar{We may want to comment part of this to make room for the 
  experiments that Lucas and Guilherme are currently running.
  Side-comparison with prior work based on paper (not actual tool) results.}
This section elaborates how \tname\ performs on two other attacks.
\textbf{Remote Code Execution (RCE).}~RCE refers to the family of
attacks where hackers exploit a network vulnerability to run arbitrary
code on a remote machine. The code to be executed could damage the
system or steal sensitive information. JBoss~\cite{jboss} is a popular
open source Java application server. This attack enables remote
execution of code on the JBoss application server.
Figure~\ref{fig:rule-jboss} shows a Suricata rule that captures a
JBoss RCE attack.  The option \CodeIn{flow: established, to\_server;}
indicates that a rule match will only occur on packets that (1) flow
on an already-established connection and that (2) flow to the server
as opposed to the client. For the content options the modifiers
\CodeIn{nocase} and \CodeIn{http\_uri} indicate, respectively, that
the matching is not case sensitive and that the string should be
located at the packet field ``Request URI''. We reproduced the
malicious JBoss attack using the traffic from a Nikto
script~\cite{nikto}. Similar attacks are also available on the
Metasploit database~\cite{metasploit}. This is a case where
\tname\ performed poorly. Of the 228 plausible rules it generated for
this attack, the ground truth was located at position 45 in the ranked
list. \textbf{Denial-of-Service.}~
The SYN Flood attack is a denial-of-service attack that exploits a
vulnerability in the tcp/ip handshake
to establish a tcp connection~\cite{cloudfare-synflood}. The handshake
works as follows in normal circumstances. First, a client sends a SYN
packet to the server, requesting a connection. Second, the server
responds with a SYN-ACK packet to the client. Third, the client
responds with an ACK message and the connection is established. Aware
of the protocol, an attacker sends multiple SYN packets to different
ports of a server, often using fake IP addresses. Then, after the
server responds with a SYN-ACK packet, the client keeps sending other
SYN packets to avoid the connection to time out. Without proper
protection, the server accepts these malicious requests and eventually
legitimate requests cannot be satisfied due to resource exhaustion.

%\vspace{-1ex}
\begin{figure}[h!]
  \lstinputlisting[language=C,numbers=none,keywords={flags,threshold}]{synflood.suricata}
%  \vspace{-2ex}
  \caption{\suri\ rule for SYN Flood attack.}
  \label{fig:synflood-example}
\end{figure}
%\vspace{-2ex}

Figure~\ref{fig:synflood-example} shows a Suricata rule that captures
this attack. This rule was obtained from the Emerging Threats Open
Ruleset~\cite{emerging-threats-open}. The relevant options in the
rule appear in bold---the option \CodeIn{flags: S,12}, identifies a
SYN packet in a tcp packet and the option \CodeIn{threshold: type
  both, track by\_dst, count 5000, seconds 5} indicates that a high
volume of such packets should be requested in a short period of
time. The other options are for documentation purposes. This is a case
where \tname\ works very well, reporting the golden rule at the top
of the ranking.

>>>>

\subsection{Discussion}
\label{sec:rce}
\label{sec:jboss}
\label{sec:content-example}
\label{sec:dos}

\subsubsection{Comparison with other tools}\Mar{Lucas and Guilherme,
  please summarize results here!}

\subsubsection{Running time}\Mar{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Add this in a journal submission in section to discuss
%% limitations...
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% \begin{figure}[t!]
%%   \lstinputlisting[language=C,numbers=none,keywords={content,flow}]{adaptor-golden-rule.suricata}
%%   \caption{\label{fig:rule-jboss}Suricata rule for JBoss Remote Code
%%     Execution.}
%% \end{figure}

%% \Mar{We may want to comment part of this to make room for the 
%%   experiments that Lucas and Guilherme are currently running.
%%   Side-comparison with prior work based on paper (not actual tool) results.}
%% This section elaborates how \tname\ performs on two other attacks.
%% \textbf{Remote Code Execution (RCE).}~RCE refers to the family of
%% attacks where hackers exploit a network vulnerability to run arbitrary
%% code on a remote machine. The code to be executed could damage the
%% system or steal sensitive information. JBoss~\cite{jboss} is a popular
%% open source Java application server. This attack enables remote
%% execution of code on the JBoss application server.
%% Figure~\ref{fig:rule-jboss} shows a Suricata rule that captures a
%% JBoss RCE attack.  The option \CodeIn{flow: established, to\_server;}
%% indicates that a rule match will only occur on packets that (1) flow
%% on an already-established connection and that (2) flow to the server
%% as opposed to the client. For the content options the modifiers
%% \CodeIn{nocase} and \CodeIn{http\_uri} indicate, respectively, that
%% the matching is not case sensitive and that the string should be
%% located at the packet field ``Request URI''. We reproduced the
%% malicious JBoss attack using the traffic from a Nikto
%% script~\cite{nikto}. Similar attacks are also available on the
%% Metasploit database~\cite{metasploit}. This is a case where
%% \tname\ performed poorly. Of the 228 plausible rules it generated for
%% this attack, the ground truth was located at position 45 in the ranked
%% list.

%% \textbf{Denial-of-Service.}~
%% The SYN Flood attack is a denial-of-service attack that exploits a
%% vulnerability in the tcp/ip handshake
%% to establish a tcp connection~\cite{cloudfare-synflood}. The handshake
%% works as follows in normal circumstances. First, a client sends a SYN
%% packet to the server, requesting a connection. Second, the server
%% responds with a SYN-ACK packet to the client. Third, the client
%% responds with an ACK message and the connection is established. Aware
%% of the protocol, an attacker sends multiple SYN packets to different
%% ports of a server, often using fake IP addresses. Then, after the
%% server responds with a SYN-ACK packet, the client keeps sending other
%% SYN packets to avoid the connection to time out. Without proper
%% protection, the server accepts these malicious requests and eventually
%% legitimate requests cannot be satisfied due to resource exhaustion.

%% %\vspace{-1ex}
%% \begin{figure}[h!]
%%   \lstinputlisting[language=C,numbers=none,keywords={flags,threshold}]{synflood.suricata}
%% %  \vspace{-2ex}
%%   \caption{\suri\ rule for SYN Flood attack.}
%%   \label{fig:synflood-example}
%% \end{figure}
%% %\vspace{-2ex}

%% Figure~\ref{fig:synflood-example} shows a Suricata rule that captures
%% this attack. This rule was obtained from the Emerging Threats Open
%% Ruleset~\cite{emerging-threats-open}. The relevant options in the
%% rule appear in bold---the option \CodeIn{flags: S,12}, identifies a
%% SYN packet in a tcp packet and the option \CodeIn{threshold: type
%%   both, track by\_dst, count 5000, seconds 5} indicates that a high
%% volume of such packets should be requested in a short period of
%% time. The other options are for documentation purposes. This is a case
%% where \tname\ works very well, reporting the golden rule at the top
%% of the ranking.

@@

Our experiments is affected by the usual threats. Considering external
validity. Our results cannot be generalized beyond the subjects we
selected. To mitigate that threat we selected a variety of attacks
with different characteristics. Unfortunately, we could not access
tools for related techniques. For that reason, we made a
side-comparison based on the common examples and results reported in
the original paper. Considering internal validity, the implementation
of \tname\ could contain errors. We mitigated that threat by carefully
inspecting code and results. Considering construct validity, the
choice of ranking to evaluate the approach may not be useful in
practice. We considered the use of ranking of the golden rule a
reasonable choice as the approach places the golden rule close to the
top in most cases.
